---
title: "Day1"
author: "Ali"
date: "2023-06-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Functions

The pieces that make a function:

```{r function recepie}
function_name <- function(param1, param2 = 20, ...){
  param1*2 # Operational space
  param1+param2 # What is returned. Alt, use return(param1+param2)
}
```

### Executing an R script

-   Interactively: `source("myscript.R")` in R console

-   Command line: `Rscript myscript.R`

-   As executable file: `path/myscript.R`

    -   if: Script is executable: chmod +x myscript.R
    -   First line in script is a hashbang e.g. #!/usr/bin/env
    -   Rscript Script's path is included in call or \$PATH

## Lab

### R Functions

```{r lab}
function_name <- function(param1, param2 = "Lucy"){
  if (param2 == "Lucy") {
    output <- paste("Lucy in the sky with ",param1,"\n",sep="")}
  else {
    output <- paste(param1," in the ocean with ",param2,"\n",sep="")}
  return(cat(output))
}
function_name("Ali")
variable_a <- function_name("Tom")
```

Now lets see the power of your function in action. Write a loop that executes your function multiple times.

```{r lab}
for (i in c("Tom","Mary","Mahesh","Henrik")){
  function_name(param1=i,param2="Brad")
  function_name(param1=i,param2="Lucy")
}
```

Now let's take a look at the special ellipsis parameter for functions. Basically the ellipsis say that a function can take additional arguments. The function should work without arguments given at the ellipsis. One great example of this is the data.frame() function.

```{r lab}
data.frame(a=c(1,2,3))
data.frame(a=c(1,2,3),b=c(4,5,6))
data.frame(a=c(1,2,3),b=c(4,5,6),d=c(7,8,9))
data.frame()
```

Another example where ellipsis are often used is to pass plot arguments into a function. Example:

```{r lab}
my_plot <- function(...){
  plot(c(1,2,3),c(1,2,3),...)
}

my_plot(main="Great title", xlab = "This will be X-label of plot", ylab="Y-label")
```

### R Scripts

```{bash lab}
Rscript myscriptfile0.R
```

### Passing and parsing arguments

```{bash lab}
./myscriptfile1.R 15
```

```{bash lab}
./myscriptfile2.R -m 5 -s 1
```

### Standard in and out

```{bash lab}
echo 100 | ./myscriptfile3.R 1> stdout.txt
echo 100 | ./myscriptfile3.R 2> stderr.txt
echo 100 | ./myscriptfile3.R &> both.txt
cat input.txt | ./myscriptfile3.R
```


## Best Coding Practices in R

-   Avoid T and F as variable names.
-   Avoid one letter variable names generally. Best way of notation in R: `snake_notation_looks_like_this`
-   Be consistent!
-   Not too long.
-   Special variable name: LETTERS, letters, month.abb, month.name, pi
-   Variable names beginning with period are hidden: .my_secret_variable 👻 will not be shown but can be accessed. You can see it with `ls(all.names=TRUE)`

#### How to write good functions

-   avoid accessing and modifying globals
    -   avoid 🚫 a \<\<- 42
    -   and 🆗 use a closure instead
    
## Lab

Task: Obscure Code
```{r}
iter_max <- 5
col_num <- 10
row_num <- 10
A <- matrix(rep(1, col_num * row_num), nrow = row_num)
for (i in 1:col_num) {
  for (j in 1:row_num) {
    if (i >= iter_max) {
      A[i, j] <- NA
    }
  }
}

# Can you improve the code more by eliminating loops or at least one of them?
```

Task: Better Formatting
```{r}
simulate_genotype <- function(q, N = 100) {
  if (length(q) == 1) {
    p <- (1 - q)
    f_gt <- c(p^2, (2 * p * q), q^2) # AA, AB, BB
  } else {
    f_gt <- q
  }
  sample(c('AA', 'AB', 'BB'), 
         size = N, 
         prob = f_gt, 
         replace = TRUE)
}
simulate_genotype(1)
simulate_genotype(0)
simulate_genotype(0.7)
```

Task: Hidden Variable
```{r}
.my_months <- c(month.abb[10:12])
.my_months <- rev(rev(month.abb)[1:3])
```

Task: Pipeline-friendly Function
```{r}
library(tidyverse)
my_filter <- function(x, threshold = 1, scalar = 5) {
  x[x >= threshold] <- NA 
  x * scalar
}

# Test:
c(-5, 5) %>% my_filter()
```

Task: Untidy Code?
```{r}
simulate_phenotype <- function(pop_params, gp_map, gtype, N) {
  pop_mean <- pop_params[1]
  pop_var <- pop_params[2]
  pheno <- rnorm(n = N, mean = pop_mean, sd = sqrt(pop_var))
  effect <- rep(0, times = length(N))
  for (gt_iter in c('AA', 'AB', 'BB')) {
    effect[gtype == gt_iter] <- rnorm(n = sum(gtype == gt_iter), 
                                      mean = gp_map[gt_iter, 'mean_eff'], 
                                      sd = sqrt(gp_map[gt_iter, 'var_eff']))
  }
  dat <- data.frame(gt = gtype, raw_pheno = pheno, effect = effect, pheno = pheno + effect)
  return(dat)
}
```

Task: Computing Variance
```{r}
sample_mean <- function(x) {
  sum(x) / length(x)
} 

sum_squared_deviations <- function(x) {
  tmp <- (x - sample_mean(x))^2
  sum(tmp)
}

std_dev <- function(x) {
  sqrt(sum_squared_deviations(x) / (length(x) - 1))
}

variance <- function(x) {
  std_dev(x)^2
}
```

Task: Writing a Wrapper Function
```{r}
randomSampleInt <- function(x, verbose, length, seed = 42) {
  if (verbose) {
    print(paste0('Generating random sample of ', length, ' integers using seed ', seed))
  }
  set.seed(seed)
  sampleInt <- sample(x = x, size = length, replace = TRUE)
  return(sampleInt)
} 

randomSampleLetter <- function(N, silent=T, lett) {
  if (!silent) {
    print(paste0('Generating random sample of ', N, ' letters.'))
  }
  sample <- sample(x = lett, size = N, replace = TRUE)
  return(sample)
}

randomSampleLetterWrapper <- function(x, verbose, length, seed = 42) {
  set.seed(seed)
  randomSampleLetter(N = length, silent = !verbose, lett = x)
}
randomSampleLetterWrapper(c("a","b","c"),F,100)
randomSampleLetterWrapper(c("a","b","c"),T,100,seed = 123)
```

Task: Customizing plot
```{r}
graphics::plot.default
my_plot <- function(x, ...) {
  plot(x, pch = 3, col = 'red', ...) 
}
```

Task: Adding Arguments to a Function
```{r}
red_plot <- function(x, y) { 
  plot(x, y, las = 1, cex.axis = .8, ...)
}

red_plot(1, 1, col = 'red', pch = 19) # Does not work.
formals(red_plot) <- c(formals(red_plot), alist(... = )) # Fix.
red_plot(1, 1, col = 'red', pch = 19) # Works!
```

Task: Using options
```{r}
.Options
options(prompt = "hello :-) > ")
.Options
options(prompt = "> ") # restoring the default
.Options
```

## Publishing with Quarto

```{r Quarto}
install.packages("quarto", dependencies = T)
```


## Debugging, Profiling, and Optimization


